---
title: "scRNA-seq Dimension Reduction with scvis"
author: "Pascale Walters"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction to scvis

```{r}
library(scvis)
library(scater)
```

`scvis` trains a probablistic parametric mapping for dimensionality reduction.

## Example Data

We'll use an example `SingleCellExperiment` object from cellassign. This object has 500 cells and 1000 genes.

```{r}
data("example_sce", package = "scvis")
example_sce
```

The counts assay contains the expressions of each gene. We'll use this for dimensionality reduction.
To reduce the noise in the data, we'll reduce to 100 dimensions. 

```{r}
example_sce <- runPCA(example_sce, ncomponents = 100, exprs_values = "counts")
```

## Training the Model

Now we can run scvis on the `SingleCellExperiment` object.
To create a new embedding, we'll run `scvis_train` on `example_sce`.

```{r}
scvis_train(example_sce, "output/", use_reducedDim = TRUE, reducedDim_name = "PCA")
```

This function performs dimension reduction on the results from `runPCA`. 
It writes the results to a .csv file in `output/`, as well as the log likelihoods, elbo results and t-SNE costs.
The reduced dimensions are also added to the `scvis` slot in the `ReducedDimensions` of `example_sce`.

The trained model is saved in `output/model/`.

## Adding New Data to an Existing Embedding

Suppose we had an existing trained model saved in `output/model/` and we wanted to add more data to it.
We can use `scvis_map` with a `SingleCellExperiment` object.

```{r}
output_dir <- file.path(find.package("scvis"), "vignette")

scvis_map(example_sce, "/home/rstudio/output1")
```

This adds the data from the `counts` assay of `example_sce` to the model that is stored in `output/model/`.
It writes the results to a .csv file in `output/`, as well as the log likelihoods.
The reduced dimensions are also added to the `scvis` slot in the `ReducedDimensions` of `example_sce`.
